Effects of Uncertainty in Heterogeneous Models

### 2017-01-24 presentation/presentation.tex
Final at the Cafe

\title{Effects of Unemployment Benefits and Uncertainty in Heterogeneous Models}

@@

%\date{Conference Name, 2013}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

@@

\begin{frame}{First Observations}
  \begin{itemize}

  \item {
  Changes in idiosyncratic risk do affect aggregate variables.
  }

  \item {
  The magnitude of impact varies for different levels of risk aversion.
  }

@@

\begin{figure}[!tbp]
  \centering
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution1}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution2}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution3}
  \end{minipage}
\end{figure}
\ \ \ \ \ \ \  UI = 0.01 \ \ \ \ \ \ \ \ \ \ \ \ UI = 0.42579 \ \ \ \ \ \ \ \ \ \ \ \ \ UI = 0.8
\end{frame}

\begin{frame}{Concluding on the distribution}

  \item {
As precautionary savings decrease, aggregate capital decreases, thus increasing the interest rate and reducing the wage. 
}
  \item {
The increase in unemployment benefits further reduces the net wage of the employed. The government has to increase income taxes to finance the higher unemployment benefits.
}
  \item {
Therefore, with higher unemployment benefits the agent's income depends more on his accumulated capital than his labor or benefits income. 
}
  \item {
The dispersion of the distribution therefore increases as benefits increase.
}

\begin{frame}{Concluding the first part}

  \item {
The presence of idiosyncratic risk leads to higher aggregate capital. 
  }
    \item {
A model with idiosyncratic risk allows for the study of distributional questions.
  } 
  \item {
Increasing unemployment benefits has distributional consequences. 
  }

@@

@@

\begin{frame}{Life-time Utility}
  \begin{itemize}
  \item {
  We calculate the life-time utility of each agent when the economy is in steady state. One time before the policy change and once after the policy change.
  }
  \item {
  Cannot compare two steady states directly. We need to take the transition into account.
  }
  \item {
  Calculating the exact transition path is troublesome. We follow the approximation by \citeauthor{sahin} \cite{sahin} which takes care of the idiosyncratic transition.
  }
    \begin{enumerate}
    \item {
    Just place the agents in the first economy with their current assets into the second economy.
    }
    \end{enumerate}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Consumption Equivalent}
  \centering{\includegraphics[scale=0.6]{cons_equiv_unemp_emp}}
\end{frame}

@@

\begin{frame}{Consumption Equivalent}
\framesubtitle{Summary}
  \begin{itemize}
  \item {
  There exists a welfare increase in the mean and in the median.
  }
  \item {
  However, there is disagreement on whether to increase or decrease the unemployment benefits.
  }
  \item {
  Careful: the consumption equivalent is biased towards poorer agents as relative increases for them are higher than for already high-consumption agents.
  }
  \end{itemize}
\end{frame}

@@

\begin{frame}{Cash Equivalent}
\framesubtitle{Summary}
  \begin{itemize}
  \item {
  There does not exist a higher welfare state in the mean, only in the median.
  }
  \item {
  If a higher welfare state existed for the mean we could redistribute wealth such that everyone would be better off. \\
  $\rightarrow$ Pareto improvement.
  }
  \item {
  One can clearly see the friction between unemployed and employed.
  }
  \end{itemize}
\end{frame}

\begin{frame}{Cash vs. Consumption Equivalent}
\centering{\includegraphics[scale=0.6]{cash_cons_equiv}}
\end{frame}

@@

  \item {
  3 transition matricies.
  }

@@

  \item {
  The introduction of aggregate uncertainty makes the distribution of wealth changing over time, which is infinite dimensional object.
  }

  \item {
  The capital distribution must be taken into account by the agents, therefore it becomes state variable.  
  }

@@

  \item {
  \citeauthor{krussel_smith97} \cite{krussel_smith97} suggest an algorithm to solve the issue by an approximation, which could be summerized as ”rational adaptive expectations”.  
  }

@@

  \item {
  Surpress Krussel - Smith algorithm such that the economy is initialy at the middle between good and bad state. i. e. $z = 1, U = 0.07$ and the probability of being employed next period is equal to the half in the 2 aggregate states. 
  }

@@

\begin{frame}{Wellfare loss of the business cycle}{Calculate lifetime utilities}
  \begin{itemize}

@@

  \item {
  Take the ergodic distribution and put it in aggregate economy framework. Solve the individual problem and caluclate the \textbf{EXPECTED} lifetime utility. 
  }

  \[\textrm{Compare: }U^1(k_i, e_i) \textrm{ and } \mathbb{E}U^2(e_i, k_i, K^{ss}, z)\]

@@

  \item {
  The median consumption equvalent is: 0.1603\%.
  }

@@

  \item {
  How to solve the individual problem? Trade off between accuracy and speed.  
  }

  \item {
  How to find the root?  
  }

@@

  \item {
  Government policies could influece the society significantly!
  }

  \item {
  There is no pareto improvement!
  }

>>>>

\title{Effects of Uncertainty in Heterogeneous Models}

@@

\date{Project Module, 25.01.2017}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

@@

@@

\begin{figure}[!tbp]
  \centering
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution1}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution2}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.325\textwidth}
    \includegraphics[width=\textwidth]{distribution3}
  \end{minipage}
\end{figure}
\ \ \ \ \ \ \  UI = 0.01 \ \ \ \ \ \ \ \ \ \ \ \ UI = 0.40 \ \ \ \ \ \ \ \ \ \ \ \ \ UI = 0.8
\end{frame}

@@

\begin{frame}{Summary}

  \item {
Idiosyncratic risk affects aggregate variables.
  }

    \item {
Increasing unemployment benefits has distributional consequences. 
  }

\item {
Welfare implications are unclear.
}

@@

\begin{frame}{Life-time Utility}
  \begin{itemize}
  \item {
  We calculate the life-time utility of each agent when the economy is in steady state. One time before the policy change and once after the policy change.
  }
  \item {
  Cannot compare two steady states directly. We need to take the transition into account.
  }
  \item {
  Calculating the exact transition path is troublesome. We follow the approximation by \citeauthor{sahin} \cite{sahin} which takes care of the idiosyncratic transition.
  }
    \begin{itemize}
    \item {
    Just place the agents in the first economy with their current assets into the second economy.
    }
    \end{itemize}
  \end{itemize}
\end{frame}

@@

@@

  \item {
  The consumption equivalent is biased towards poorer agents as relative increases for them are higher than for already-high consumption agents. Therefore, we need another measure.
  }
  \end{itemize}
\end{frame}

@@

@@

@@

  \item {
  The introduction of aggregate uncertainty makes the distribution of wealth changing over time, which is an infinite dimensional object.
  }

  \item {
  The capital distribution must be taken into account by the agents, therefore it becomes a state variable.  
  }

@@

  \item {
  \citeauthor{krussel_smith97} \cite{krussel_smith97} suggested an algorithm to solve the issue by an approximation.  
  }

@@

  \item {
  Surpress Krussel - Smith algorithm such that the economy is initialy at the middle between good and bad state.
  }

@@

\begin{frame}{Welfare loss of the business cycle}{Calculate lifetime utilities}
  \begin{itemize}

@@

  \item {
  Take the ergodic distribution and put it in aggregate economy framework. Solve the individual problem and calculate the \textbf{EXPECTED} lifetime utility. 
  }

  \[\textrm{Compare: }U^1(k_i, e_i) \textrm{ and } \mathbb{E}U^2(k_i, e_i, K^{ss}, z)\]

@@

  \item {
  The median consumption equivalent is: 0.1603\%.
  }

@@

  \item {
  How to solve the individual problem?  
  }

@@

  \item {
  Government policies could influece the economy significantly!
  }

  \item {
  There is no Pareto improvement!
  }
